Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-10-24 paper-new/discussion.tex
discussion stuff

Additionally, while we have focused on a single semantics in this work, there are a
variety of other cast calculi that satisfy graduality but for which one or more
equational axioms fail to hold. \eric{Max, please add examples}
We expect that our development can be adapted to other cast semantics to prove
graduality and equational reasoning principles (to the extent that the
principles hold in those cast semantics).

We also plan to use our framework to investigate the performance of
gradually-typed languages. Although we have focused on proving extensional
properties in this work, the intensional nature of our semantics can be viewed
as a form of \emph{cost semantics}, where a runtime cost is incurred from
inspecting the dynamic type. This could be used to verify that cast
optimizations such as space efficient implementations
\cite{herman-tomb-flanagan-2010} are not only extensionally correct, but also
have lower quantitative cost.

>>>>

This work has focused on the ``Natural'' semantics of casts , which
validate the full call-by-value $\eta$ laws for types, but other cast
semantics have been proposed such as eager
(\cite{herman-tomb-flanagan-2010}) and transient (\cite{transient})
cast semantics which trade off certain $\eta$ equalities for other
benefits such as early error detection or reduced runtime
overhead\cite{deepandshallowtypes,newlicataahmed19}. It would be a
good test of the generality of our semantic framework to see if it can
model these alternative semantics by adjusting the semantics of types
and casts, and providing proofs of weakened $\eta$ principles.

Additionally, while our focus in this work has been on reasoning up to
weak bisimilarity, the presence of explicit step counting in the model
could be viewed as a form of \emph{cost semantics}, where a runtime
cost is incurred from inspecting the dynamic type. This could possibly
be used to verify that cast optimizations such as space efficient
implementations \cite{herman-tomb-flanagan-2010} are not only
extensionally correct, but have a lower abstract cost.
